---
title: "Sample R scripts for Consultant Data Analytics, GMDAC Job Offer"
author: "Hedi Ben Mustapha"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    highlight: null
    css: style.css
    df_print: paged
    mathjax: local
    self_contained: false
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

```{r}
load("../.RData")
```

# FOOD SECURITY


## FCS (Food Consumption Score)
```{r}
p1 
```

## rCSI (reduced Coping Strategy Index)
```{r}
p2 
```


## LCSI (Livelihood Coping Strategy Index)
```{r}
p3
```


## Food Sources
```{r}
p4

```


# WATER, SANITATION & HYGIENE (WASH) {.tabset .tabset-fade .tabset-pills}


## Water sources
```{r}
p5
```


## Access for public water network
```{r}
p6
```


# CASH & MARKETS

## Distribution of income

```{r}
p7
```

## Distribution of income VS Distribution of expenditure
```{r}
p8
```

## Income & revenue data {.tabset .tabset-fade .tabset-pills}

### Median Income
```{r}
income_dt

```

### Median expenditure
```{r}
expenditure_dt

```

