---
title: "Sample R scripts for Consultant Data Analytics, GMDAC Job Offer"
author: "Hedi Ben Mustapha"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    highlight: null
    css: style.css
    df_print: paged
    mathjax: local
    self_contained: false
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

```{r functions}
#install only packages not available in the library
install_new_packages <- function(list.of.packages){
  new.packages <- list.of.packages[!(list.of.packages %in%                                      installed.packages()[,"Package"])]
  if(length(new.packages)) install.packages(new.packages) 
}

#sanitize dataframe headers
to_alphanumeric_lowercase <- function(x){
  tolower(gsub("[^a-zA-Z0-9_]", "\\.", x))
}

#read csv files
import_csv <- function(path_tofile){
  assertthat::assert_that(grepl(x = path_tofile, pattern = ".csv$"), 
                          msg = "file must end with '.csv' (..and actually be a .csv file)")
  df <- data.table::fread(file = path_tofile, 
                          stringsAsFactors = F,
                          na.strings = c("","na","n/a","NA",NA)) %>%                                                     as.data.frame
  colnames(df) <- to_alphanumeric_lowercase(colnames(df))
  return(df)
}

#Creates a weighting function from sampling frame
create_weights <- function(sampling.frame, data.stratum.column, 
                           sampling.frame.population.column = "population", 
                           sampling.frame.stratum.column = "strata.names", data = NULL){
  surveyweights::weighting_fun_from_samplingframe(
    sampling.frame =  sampling.frame, 
    data.stratum.column = data.stratum.column, 
sampling.frame.population.column = sampling.frame.population.column, 
sampling.frame.stratum.column = sampling.frame.stratum.column, 
data = data)
}

plot_animated_ggplot <- function(summary_object, categories, numbers, states,plot_title){
  categories <- sym(categories)
  numbers <- sym(numbers)
  states <- sym(states)
  ggplot(summary_object, aes(x = !!categories, 
                        y = !!numbers * 100, fill = !!categories)) +
  geom_col(position = 'dodge') + 
  labs(title = plot_title,
       x = NULL,
       y = NULL) +
  theme_minimal() + 
  theme(
    legend.position = 'none',      
    plot.title = element_text(hjust = 0.5),
    text = element_text('Arial Narrow'),
    strip.text = element_text(face = 'bold', hjust = 0),
    panel.grid.minor = element_blank()
  ) +
  geom_text(aes(label = paste(round(!!numbers * 100), "%")),
            position=position_dodge(width=0.9),
            vjust = -0.5) +
  transition_states(states = !!states) +
  enter_grow() + 
  exit_fade()
}

plot_flipped_ggplot <- function(summary_object, categories, numbers, plot_title){
  categories <- sym(categories)
  numbers <- sym(numbers)
  ggplot(summary_object, aes(x = reorder(!!categories, !!numbers), 
                               y = !!numbers * 100, fill = !!categories)) +
  geom_col(position = 'dodge') + 
  labs(title = plot_title,
       x = NULL,
       y = NULL) +
  theme_minimal() + 
  theme(
    legend.position = 'none',      
    plot.title = element_text(hjust = 0.5),
    text = element_text('Arial Narrow'),
    strip.text = element_text(face = 'bold', hjust = 0),
    panel.grid.minor = element_blank()
  ) +
  geom_text(aes(label = paste(round(!!numbers * 100), "%")),
            position=position_dodge(width=0.9),
            hjust = -0.5) +
  ylim(0, 100) +
  coord_flip()
}

plot_animated_flipped_ggplot <- function(summary_object, categories, numbers, states, plot_title){
  categories <- sym(categories)
  numbers <- sym(numbers)
  states <- sym(states)
  ggplot(summary_object, aes(x = reorder(!!categories, !!numbers), 
                             y = !!numbers * 100, fill = !!categories)) +
    geom_col(position = 'dodge') + 
    labs(title = plot_title,
         x = NULL,
         y = NULL) +
    theme_minimal() + 
    theme(
      legend.position = 'none',      
      plot.title = element_text(hjust = 0.5),
      text = element_text('Arial Narrow'),
      strip.text = element_text(face = 'bold', hjust = 0),
      panel.grid.minor = element_blank()
    ) +
    geom_text(aes(label = paste(round(!!numbers * 100), "%")),
              position=position_dodge(width=0.9),
              hjust = -0.5) +
    ylim(0, 100) +
    coord_flip() +
    transition_states(states = !!states) +
    enter_grow() + 
    exit_fade()
}

f <- function(x) {
  r <- quantile(x, probs = c(0.00, 0.25, 0.5, 0.75, 1))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r}

#Map results to an output template
render_report_rmd <- function(x, dir, filename) {
  
  template <- "code/report.Rmd"
  render_environment <- new.env()
  
  render_environment$x <- x
  rmarkdown::render(
    template,
    output_file = filename,
    output_dir = dir,
    intermediates_dir = dir,
    envir = render_environment,
    knit_root_dir = getwd()
  )
  
}
```

```{r dependencies, eval=TRUE, results= 'hide'}
list.of.packages <- c(
  'assertthat',
  'data.table',
  'surveyweights',
  'stringr',
  'tidyverse',
  'srvyr',
  'gganimate',
  'ggthemes',
  'plotly',
  'extrafont',
  'DT',
  'magick',
  'magrittr'
)

#Install new packages
install_new_packages(list.of.packages)

library(assertthat)
library(data.table)
library(surveyweights)
library(stringr)
library(tidyverse)
library(srvyr)
library(gganimate)
library(ggthemes)
library(plotly)
library(extrafont)
library(DT)
library(magick)
library(magrittr)
fonts()
fonttable()
```

```{r data_preparation}
# Load dataset
data <- import_csv(path_tofile = "../input/data.csv")

#Load sampling frame
sampling_frame <- import_csv(path_tofile = "../input/sampling_frame.csv")

# adding new indicators
data <- data %>% mutate(
  strata.names = str_c(mantika_label,displacement_status,sep = "_"),
  fcs = (cereals * 2) + (legumes * 3) + veggies + fruits + (meat * 4) + (dairy * 4) + (fats * 0.5) + (sugar * 0.5),
  fcs_category = case_when(
    fcs <= 28 ~ "poor",
    fcs > 28 & fcs <= 42 ~ "borderline",
    fcs > 42 ~ "acceptable"
  ),
  rcsi = less_expensive_quality + (borrow_relatives * 2) + reduce_number_meals + (reduce_adult * 3) + shrink_meals,
  rcsi_category = case_when(
    rcsi <= 3 ~ "low",
    rcsi > 3 & rcsi <= 9 ~ "medium",
    rcsi > 9 ~ "high"
  ),
  total_income = rowSums(select(.,gvt_salary,gvt_social_benefits,
                                non_gvt_salary,casual_labour,
                                own_business_income,
                                remittances,
                                family_support,
                                humanitarian_assistance,
                                zakat,income_other) 
                         , na.rm=T),
  total_income = ifelse(total_income ==0, NA,total_income),
  total_expenditures = rowSums(select(.,food_expenditure,
                                      rent_expenditure,
                                      shelter_maintenance_expenditure,
                                      water_expenditure,
                                      nfi_expenditure,
                                      utilities_expenditure,
                                      fuel_expenditure,
                                      health_related_expenditure,
                                      education_related_expenditure
                                      ,
                                      transportation_expenditure,
                                      mobile_phone_credit_expenditure,
                                      productive_assets_expenditure,
                                      debt_repayment_expenditure,
                                      other_expenditure),
                               na.rm=T),
  total_expenditures = ifelse(total_expenditures == 0, NA, total_expenditures),
  lcsi_stress = rowSums(select(., sold_nonproductive_hh_assets,
                                  spent_savings,
                                  borrowed_purchased_oncredit_food,
                                  reduced_expenditures_essential_nfi) %>%
                                 mutate_all(~ .x %in% c("already_exhausted_this_coping_strategy", "yes")),
                               na.rm = T),
  lcsi_crisis = rowSums(select(.,sold_productive_hh_assets,
                                borrowed_money,
                                reduced_expenditures_health_education,
                                took_additional_job,
                                child_dropped_school,
                                delayed_skipped_rent) %>%
                                 mutate_all(~ .x %in% c("already_exhausted_this_coping_strategy", "yes")),
                               na.rm = T),
  lcsi_emergency = rowSums(select(.,begging,
                                  adult_accepting_degrading_illegal_work,
                                  minor_accepting_degrading_illegal_work,
                                  child_marriage) %>%
                                    mutate_all(~ .x %in% c("already_exhausted_this_coping_strategy", "yes")),
                                  na.rm = T),
  lcsi = case_when(
    lcsi_emergency > 0 ~ "emergency",
    lcsi_crisis > 0 ~ "crisis",
    lcsi_stress > 0 ~ "stress",
    lcsi_emergency + lcsi_crisis + lcsi_stress == 0 ~ "none"
  )
)

#create weighting fct
weights <- create_weights(sampling.frame = sampling_frame,
                          data.stratum.column = "strata.names",
                          sampling.frame.population.column = "population",
                          sampling.frame.stratum.column = "strata.names",
                          data = data)

data <- data %>% mutate(
  strata.weights = weights(data)
)

```

```{r data_analysis}
# create survey design
design <- data %>% as_survey(weights = c(strata.weights))

fcs_summary <- design %>% 
  group_by(displacement_status,fcs_category) %>% 
  summarize(numbers = survey_mean())

rcsi_summary <- design %>% 
  group_by(displacement_status,rcsi_category) %>% 
  summarize(numbers = survey_mean())


lcsi_summary <- design %>% 
  group_by(displacement_status,lcsi) %>% 
  summarize(numbers = survey_mean(na.rm =T))

foodsource_summary <- design %>%  
  summarize(Market_purchased_cash = survey_mean(food_source.market_cash),
            Market_purchased_cheque = survey_mean(food_source.market_cheque),
            Market_purchased_credit = survey_mean(food_source.market_credit)) %>%
  pivot_longer(c(Market_purchased_cash,
                 Market_purchased_cheque,
                 Market_purchased_credit))


watersource_summary <- design %>% 
  group_by(primary_drinkingwater_source) %>%
  summarize(numbers = survey_mean()) %>% 
  arrange(-numbers)

access_publicwater_summary <- design %>% 
  group_by(displacement_status, access_public_water_network) %>%
  summarize(numbers = survey_mean()) %>% 
  arrange(-numbers)

 income_summary <- design %>% 
   group_by(displacement_status) %>%
  summarize(
    numbers = survey_median(total_income, na.rm = T, vartype = NULL),
    income = survey_quantile(total_income, quantile = c(0,0.25, 0.75,1), vartype = NULL, na.rm = T)) %>% 
   mutate(level = "income")
 
 income_summary$displacement_status <- with(income_summary,
                                            reorder(income_summary$displacement_status,
        numbers,function(x) - median(x, na.rm=TRUE)))
 
 income_summary_strata <- design %>% group_by(mantika_label,displacement_status) %>%
   summarize(
     numbers = survey_median(total_income, na.rm = T, vartype = NULL)
   )
 
 expenditure_summary <- design %>% group_by(displacement_status) %>%
   summarize(
     numbers = survey_median(total_expenditures, na.rm = T, vartype = NULL),
     expenditure = survey_quantile(total_expenditures, quantile = c(0,0.25, 0.75,1), vartype = NULL, na.rm = T)
   ) %>% mutate(
     level = "expenditure"
   )
 
  expenditure_summary_strata <- design %>% group_by(mantika_label,displacement_status) %>%
   summarize(
     numbers = survey_median(total_expenditures, na.rm = T, vartype = NULL)
   )
 
 income_expenditure <- bind_rows(income_summary,expenditure_summary)

```

# FOOD SECURITY


## FCS (Food Consumption Score)
```{r}
plot_animated_ggplot(fcs_summary,
                     "fcs_category",
                     "numbers",
                     "displacement_status",
                     "% of HHs with an acceptable, borderline or poor 
                     FCS (Food Consumption Score) for {closest_state}") 
```

## rCSI (reduced Coping Strategy Index)
```{r}
plot_animated_ggplot(rcsi_summary,
                     "rcsi_category",
                     "numbers",
                     "displacement_status",
                     "% of HHs with low, medium or high rCSI 
                    (reduced Coping Strategy Index) for {closest_state}")
```


## LCSI (Livelihood Coping Strategy Index)
```{r}
plot_animated_ggplot(lcsi_summary,
                     "lcsi",
                     "numbers",
                     "displacement_status",
                     "% of HHs with a stress, crisis or emergency LCSI 
  (Livelihood Coping Strategy Index) for {closest_state}")
```


## Food Sources
```{r}
plot_flipped_ggplot(foodsource_summary,
                    "name",
                    "value",
                    "Top 3 sources from which households reported acquiring food")

```


# WATER, SANITATION & HYGIENE (WASH) {.tabset .tabset-fade .tabset-pills}


## Water sources
```{r}
plot_flipped_ggplot(watersource_summary,
                    "primary_drinkingwater_source",
                    "numbers",
                    "Main reported sources of drinking water")
```


## Access for public water network
```{r}
plot_animated_flipped_ggplot(access_publicwater_summary,
                             "access_public_water_network","numbers",
                              "displacement_status",
                              "access to the water from the public network in the last 7 days for {closest_state}")
```


# CASH & MARKETS

## Distribution of income

```{r}
ggplot(income_summary, aes(x=displacement_status, 
                            ymin = income_q00, 
                            lower = income_q25, 
                            middle = numbers, 
                            upper = income_q75, 
                            ymax = income_q100,
                            width = 0.7, 
                            fill = displacement_status)) +
   geom_boxplot(stat = "identity") +
   theme_bw() +
   theme(legend.position = 'none',
         plot.title = element_text(hjust = 0.5),
         text = element_text('Arial Narrow'),
         panel.grid.major.x = element_blank(),
         panel.grid.major.y = element_line(size=.1, color="grey85"),
         panel.grid.minor = element_blank(),
         axis.text.x = element_text(angle = 0, size = 10.2, hjust = 0.5)) +
   labs(title = "Median total income by displacement groups",
        x = NULL,
        y = " Income (LYD)") +
   geom_text(data = income_summary, 
             aes(y = numbers, label = round(numbers,digits = 0)),
             size = 3.52, hjust = -2.57, vjust =  0.6)
```

## Distribution of income VS Distribution of expenditure
```{r}
plot_ly(income_expenditure, 
        x = ~displacement_status, 
        y = ~numbers, type = 'bar',
        color = ~level) %>%
   layout(title = "Median total income VS Median Total expenditure",               yaxis = list(title = "LYD"), 
          xaxis = list(title = ""), 
          barmode = 'group')
```

## Income & revenue data {.tabset .tabset-fade .tabset-pills}

### Median Income
```{r}
   income_summary_strata %>% 
   dplyr::rename(Medians = numbers) %>% 
   datatable()

```

### Median expenditure
```{r}
   expenditure_summary_strata %>% 
   dplyr::rename(Medians = numbers) %>% 
   datatable()

```

